Supporting information - Bayesian regression facilitates quantitative modelling of cell metabolism

### 2024-01-09 paper/supporting_information.tex
editorial changes

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

@@

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Supporting information},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Supporting information}
\author{}
\date{}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, interior hidden, sharp corners, breakable, enhanced]}{\end{tcolorbox}}\fi

@@

Maud inputs are structured directories, somewhat inspired by the PEtab
format (Schmiester et al. 2021). A Maud input directory must contain a
toml (Preston-Werner, Tom and Gedam, Pradyun 2020) file called
\texttt{config.toml} which gives the input a name, configures how Maud
will be run and tells Maud where to find the other files, allowing these
to have custom names. It must also include a file containing a kinetic
model definition, a file specifying information about parameters and a
file with information experiments. The required structure of these files
is documented at
\url{https://maud-metabolic-models.readthedocs.io/en/latest/inputting.html}.
The input is validated against a Pydantic (Pydantic developers 2022)
data model.

We chose to implement a custom input format despite the existence of
standard formats in similar areas, including SBML (Keating et al. 2020)
and PEtab (Schmiester et al. 2021). This choice was partly motivated by
the need to ensure flexibility as Maud was developed, but there are also
features of SBML and PEtab that make them structurally unsuitable in
this context. Our requirements for an input format included that it be
mathematics-free, so that all mathematical details are encapsulated in
source code, and that it has a detailed, verifiable structure. These
requirements made toml more attractive than SBML: toml is easier for
humans to read and edit and can straightforwardly be validated using
tools like Pydantic. Further, an SBML representation of our desired
input would not contain differential equations. It would therefore not
be interoperable with most SBML targeting software, which typically
assumes that differential equations are available and does not know
about Maud's structure.

@@

\hypertarget{tbl-params}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4722}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2778}}@{}}
\caption{\label{tbl-params}Parameters of Maud's statistical
model}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Parameter
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Modelled quantity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Dimensions
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Parameter
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Modelled quantity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Dimensions
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(\Delta_fG\) & Formation energy & metabolites \\
\(k_M\) & Michaelis Menten constants & Substrates of all
enzyme/reactions and products of reversible enzyme/reactions \\
\(k_I\) & Inhibition constants & Inhibiting metabolite/compartments of
enzyme/reactions exhibiting competitive inhibition \\
\(k_{cat}\) & Rate constants & Enzyme/reactions \\
\(L_0\) & Transfer constants & Allosteric interactions \\
\(e_T\) & T dissociation constants & Modifying metabolites of allosteric
inhibitions \\
\(e_R\) & R dissociation constants & Modifying metabolites of allosteric
activations \\
\(k_{cat\, pme}\) & Rate constants of phosphorylation modifying enzymes
& Phosphorylation modifying enzymes \\
\(v_{drain}\) & Drain fluxes & Drains, experiments \\
\(Enzyme\) & Enzyme concentrations & Enzymes, experiments \\
\(C_{unbalanced}\) & Unbalanced metabolite/compartment concentrations &
Unbalanced metabolite/compartments, experiments \\
\(C_{pme}\) & Phosphorylation modifying enzyme concentrations &
Phosphorylation modifying enzymes, experiments \\
\(\psi\) & Membrane potentials & Experiments \\
\end{longtable}

@@

The term \(Saturation\) in equation \eqref{eq-decomposition} is a vector
of non-negative real numbers representing, for each reaction, the
fraction of enzyme that is saturated, i.e.~bound to one of the
reaction's substrates. To describe saturation we use equation
\eqref{eq-saturation}, which is taken from Liebermeister, Uhlendorf, and
Klipp (2010) and Noor et al. (2013). Additionally, this term captures
competitive inhibition: as competitive inhibitor concentration
increases, the saturation denominator increases, effectively decreasing
the saturation of the substrate on the total enzyme pool. Conversely, as
the substrate concentration increases this term approaches 1.

@@

The term \(Allostery\) in equation \eqref{eq-decomposition} is a vector
of non-negative numbers describing the effect of allosteric regulation
on each reaction. Allosteric regulation happens when binding to a
certain molecule changes an enzyme's shape in a way that changes its
catalytic behaviour. We use equation \eqref{eq-allostery} to describe
this phenomenon, following the generalised MWC approach described in
Monod, Wyman, and Changeux (1965), Changeux (2013), Popova and Sel'kov
(1975) and Popova and Sel'kov (1979).

@@

Starting with the model in Saa and Nielsen (2016), we extracted values
for enzyme concentrations, boundary conditions and fluxes. We used these
values to generate MCMC samples using Maud using the priors specified in
section Section~\ref{sec-methionine-priors}. When this was finished, we
selected one sample with relatively high log probability to use as a
ground truth in our case study. These parameter values are shown below
in table Table~\ref{tbl-case-study-params}. We manually inspected the
parameter values to screen for any obviously implausible values; we did
not find any of these.

@@

\caption{\label{tbl-case-study-params}Parameter specification, marginal
prior distributions and true parameter values used in our case
study.}\tabularnewline

@@

\(\Delta_fG\) parameters for most metabolites were fixed; those that
were modelled as unknown had a multivariate normal prior distribution
derived from eQuilibrator (Beber et al. 2021).

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-beberEQuilibratorPlatformEstimation2021}{}}%
Beber, Moritz E., Mattia G. Gollub, Dana Mozaffari, Kevin M. Shebek, and
Elad Noor. 2021. {``{eQuilibrator} 3.0 -- a Platform for the Estimation
of Thermodynamic Constants.''} \emph{arXiv:2103.00621 {[}q-Bio{]}},
February. \url{http://arxiv.org/abs/2103.00621}.

\leavevmode\vadjust pre{\hypertarget{ref-changeux_2013}{}}%
Changeux, Jean-Pierre. 2013. {``50 Years of Allosteric Interactions: The
Twists and Turns of the Models.''} \emph{Nature Reviews. Molecular Cell
Biology} 14 (12): 819--29. \url{https://doi.org/10.1038/nrm3695}.

\leavevmode\vadjust pre{\hypertarget{ref-keatingSBMLLevelExtensible2020}{}}%
Keating, Sarah M, Dagmar Waltemath, Matthias König, Fengkai Zhang,
Andreas Dräger, Claudine Chaouiya, Frank T Bergmann, et al. 2020.
{``{\textsc{SBML}} {Level} 3: An Extensible Format for the Exchange and
Reuse of Biological Models.''} \emph{Molecular Systems Biology} 16 (8):
e9110. \url{https://doi.org/10.15252/msb.20199110}.

\leavevmode\vadjust pre{\hypertarget{ref-liebermeister_modular_2010}{}}%
Liebermeister, Wolfram, Jannis Uhlendorf, and Edda Klipp. 2010.
{``Modular Rate Laws for Enzymatic Reactions: Thermodynamics,
Elasticities and Implementation.''} \emph{Bioinformatics} 26 (12):
1528--34. \url{https://doi.org/10.1093/bioinformatics/btq141}.

\leavevmode\vadjust pre{\hypertarget{ref-monod_nature_1965}{}}%
Monod, J, J Wyman, and J P Changeux. 1965. {``On the Nature of
Allosteric Transitions: A Plausible Model.''} \emph{Journal of Molecular
Biology} 12 (May): 88--118.
\url{https://doi.org/10.1016/S0022-2836(65)80285-6}.

\leavevmode\vadjust pre{\hypertarget{ref-noor_note_2013}{}}%
Noor, Elad, Avi Flamholz, Wolfram Liebermeister, Arren Bar-Even, and Ron
Milo. 2013. {``A Note on the Kinetics of Enzyme Action: {A}
Decomposition That Highlights Thermodynamic Effects.''} \emph{FEBS
Letters}, A century of {Michaelis} - {Menten} kinetics, 587 (17):
2772--77. \url{https://doi.org/10.1016/j.febslet.2013.07.028}.

\leavevmode\vadjust pre{\hypertarget{ref-popova_generalization_1975}{}}%
Popova, S V, and E E Sel'kov. 1975. {``Generalization of the Model by
{Monod}, {Wyman} and {Changeux} for the Case of a Reversible
Monosubtrate Reaction {SR},{TP}.''} \emph{FEBS Letters} 53 (3): 269--73.
\url{https://doi.org/10.1016/0014-5793(75)80034-2}.

\leavevmode\vadjust pre{\hypertarget{ref-popova_description_1979}{}}%
---------. 1979. {``{[}{Description} of the Kinetics of the Two
Substrate Reactions {S1}+ {S2} Goes to and Comes from {S3}+{S4} by a
Generalized {Monod}, { Wyman}, {Changeux} Model{]}.''}
\emph{Molekuliarnaia Biologiia} 13 (1): 129--39.
\url{https://www.ncbi.nlm.nih.gov/pubmed/156878}.

\leavevmode\vadjust pre{\hypertarget{ref-preston-wernertomandgedampradyunTOMLSpecification0rc2020}{}}%
Preston-Werner, Tom and Gedam, Pradyun. 2020. {``{TOML Specification}
1.0.0-Rc.1.''} \url{https://toml.io/en/v1.0.0-rc.1/}.

\leavevmode\vadjust pre{\hypertarget{ref-pydanticdevelopersPydantic2022}{}}%
Pydantic developers. 2022. {``Pydantic.''}
\url{https://pypi.org/project/pydantic/}.

\leavevmode\vadjust pre{\hypertarget{ref-saa_construction_2016}{}}%
Saa, Pedro A, and Lars K Nielsen. 2016. {``Construction of Feasible and
Accurate Kinetic Models of Metabolism: {A} {Bayesian} Approach.''}
\emph{Scientific Reports} 6 (July): 29635.
\url{https://doi.org/10.1038/srep29635}.

\leavevmode\vadjust pre{\hypertarget{ref-SchmiesterSch2021}{}}%
Schmiester, Leonard, Yannik Schälte, Frank T. Bergmann, Tacio Camba,
Erika Dudkin, Janine Egert, Fabian Frö hlich, et al. 2021.
{``{PEtab}\textemdash{{Interoperable}} Specification of Parameter
Estimation Problems in Systems Biology.''} \emph{PLOS Computational
Biology} 17 (1): 1--10.
\url{https://doi.org/10.1371/journal.pcbi.1008646}.

\end{CSLReferences}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[journal=asbcd6,manuscript=article,layout=traditional]{achemso}
\usepackage[version=3]{mhchem}
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}

@@

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\renewcommand{\thepage}{S\arabic{page}}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Supporting information - Bayesian regression facilitates quantitative modelling of cell metabolism},
  pdfauthor={Teddy Groves; Nicholas Luke Cowie; Lars Keld Nielsen},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\author{Teddy Groves}
\affiliation{The Novo Nordisk Foundation Center for
Biosustainability, DTU, Kongens Lyngby 2800, Denmark}

\altaffiliation{These authors contributed equally to this research.}

\email{tedgro@biosustain.dtu.dk}
\author{Nicholas Luke Cowie}
\affiliation{The Novo Nordisk Foundation Center for
Biosustainability, DTU, Kongens Lyngby 2800, Denmark}

\altaffiliation{These authors contributed equally to this research.}

\author{Lars Keld Nielsen}
\affiliation{The Novo Nordisk Foundation Center for
Biosustainability, DTU, Kongens Lyngby 2800, Denmark}
\affiliation{Australian Institute for Bioengineering and Nanotechnology
(AIBN), The University of Queensland, St Lucia 4067, Australia}

\title[]{Supporting information - Bayesian regression facilitates
quantitative modelling of cell metabolism}
\makeatletter
\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, frame hidden]}{\end{tcolorbox}}\fi

\newpage{}

@@

Maud inputs are structured directories, somewhat inspired by the PEtab
format \citep{SchmiesterSch2021}. A Maud input directory must contain a
toml \citep{preston-wernertomandgedampradyunTOMLSpecification0rc2020}
file called \texttt{config.toml} which gives the input a name,
configures how Maud will be run and tells Maud where to find the other
files, allowing these to have custom names. It must also include a file
containing a kinetic model definition, a file specifying information
about parameters and a file with information experiments. The required
structure of these files is documented at
\url{https://maud-metabolic-models.readthedocs.io/en/latest/inputting.html}.
The input is validated against a Pydantic
\citep{pydanticdevelopersPydantic2022} data model.

We chose to implement a custom input format despite the existence of
standard formats in similar areas, including SBML
\citep{keatingSBMLLevelExtensible2020} and PEtab
\citep{SchmiesterSch2021}. This choice was partly motivated by the need
to ensure flexibility as Maud was developed, but there are also features
of SBML and PEtab that make them structurally unsuitable in this
context. Our requirements for an input format included that it be
mathematics-free, so that all mathematical details are encapsulated in
source code, and that it has a detailed, verifiable structure. These
requirements made toml more attractive than SBML: toml is easier for
humans to read and edit and can straightforwardly be validated using
tools like Pydantic. Further, an SBML representation of our desired
input would not contain differential equations. It would therefore not
be interoperable with most SBML targeting software, which typically
assumes that differential equations are available and does not know
about Maud's structure.

@@

\hypertarget{tbl-params}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4722}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2778}}@{}}
\caption{\label{tbl-params}Table S1 -- Parameters of Maud's statistical
model}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Parameter
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Modelled quantity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Dimensions
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Parameter
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Modelled quantity
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Dimensions
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(\Delta_fG\) & Formation energy & metabolites \\
\(k_M\) & Michaelis Menten constants & Substrates of all
enzyme/reactions and products of reversible enzyme/reactions \\
\(k_I\) & Inhibition constants & Inhibiting metabolite/compartments of
enzyme/reactions exhibiting competitive inhibition \\
\(k_{cat}\) & Rate constants & Enzyme/reactions \\
\(L_0\) & Transfer constants & Allosteric interactions \\
\(e_T\) & T dissociation constants & Modifying metabolites of allosteric
inhibitions \\
\(e_R\) & R dissociation constants & Modifying metabolites of allosteric
activations \\
\(k_{cat\, pme}\) & Rate constants of phosphorylation modifying enzymes
& Phosphorylation modifying enzymes \\
\(v_{drain}\) & Drain fluxes & Drains, experiments \\
\(Enzyme\) & Enzyme concentrations & Enzymes, experiments \\
\(C_{unbalanced}\) & Unbalanced metabolite/compartment concentrations &
Unbalanced metabolite/compartments, experiments \\
\(C_{pme}\) & Phosphorylation modifying enzyme concentrations &
Phosphorylation modifying enzymes, experiments \\
\(\psi\) & Membrane potentials & Experiments \\
\end{longtable}

@@

The term \(Saturation\) in equation \eqref{eq-decomposition} is a vector
of non-negative real numbers representing, for each reaction, the
fraction of enzyme that is saturated, i.e.~bound to one of the
reaction's substrates. To describe saturation we use equation
\eqref{eq-saturation}, which is taken from
\citet{liebermeister_modular_2010} and \citet{noor_note_2013}.
Additionally, this term captures competitive inhibition: as competitive
inhibitor concentration increases, the saturation denominator increases,
effectively decreasing the saturation of the substrate on the total
enzyme pool. Conversely, as the substrate concentration increases this
term approaches 1.

@@

The term \(Allostery\) in equation \eqref{eq-decomposition} is a vector
of non-negative numbers describing the effect of allosteric regulation
on each reaction. Allosteric regulation happens when binding to a
certain molecule changes an enzyme's shape in a way that changes its
catalytic behaviour. We use equation \eqref{eq-allostery} to describe
this phenomenon, following the generalised MWC approach described in
\citet{monod_nature_1965}, \citet{changeux_2013},
\citet{popova_generalization_1975} and \citet{popova_description_1979}.

@@

Starting with the model in \citet{saa_construction_2016}, we extracted
values for enzyme concentrations, boundary conditions and fluxes. We
used these values to generate MCMC samples using Maud using the priors
specified in section Section~\ref{sec-methionine-priors}. When this was
finished, we selected one sample with relatively high log probability to
use as a ground truth in our case study. These parameter values are
shown below in table Table~\ref{tbl-case-study-params}. We manually
inspected the parameter values to screen for any obviously implausible
values; we did not find any of these.

@@

\caption{\label{tbl-case-study-params}Table S2 -- Parameter
specification, marginal prior distributions and true parameter values
used in our case study.}\tabularnewline

@@

\(\Delta_fG\) parameters for most metabolites were fixed; those that
were modelled as unknown had a multivariate normal prior distribution
derived from eQuilibrator
\citep{beberEQuilibratorPlatformEstimation2021}.

@@

\renewcommand{\bibsection}{}
\bibliography{bibliography.bib}
